A Mechanised Library for Generic (Co)Programming

### 2024-01-26 paper/paper.tex
Coq implementation of the Mutu

\hypersetup{
  linktocpage,
  colorlinks,
  citecolor=BlueViolet,
  filecolor=red,
  linkcolor=Blue,
  urlcolor=BrickRed
}
\newcommand{\mpav}[1]{\textcolor{red}{\textsc{Marco}: #1}}
\newcommand{\dcas}[1]{\textcolor{red}{\textsc{David}: #1}}
\newcommand{\proofcomment}[1]{\text{\{ #1 \}}}
\newenvironment{proofof}[1] {\begin{proof}[Proof of {#1}]}{\end{proof}}
\newcommand{\eqdef}{\stackrel{\mathrm{\Delta}}{=}}
\newcommand{\bnfeq}{\mathrel{::=}}
\newcommand{\defeq}{\triangleq}
\newcommand{\rul}[3]{\frac{#2}{#3}\;  {\textrulelabel{#1}}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\jud}[3]{#1 \vdash #2 : #3}
\newcommand{\bden}[1]{\llparenthesis#1 \rrparenthesis}
\newcommand{\bigslant}[2]{{\raisebox{.2em}{$#1$}\left/\raisebox{-.2em}{$#2$}\right.}}
\newcommand{\quotient}[2]{\bigslant{#1}{#2}}
\newcommand{\curry}{\Lambda}
\newcommand{\uncurry}{\begin{sideways}\begin{sideways}$\Lambda$\end{sideways}\end{sideways}}
\newcommand{\Bool}{\mathbb{B}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Nat}{\N}
\newcommand{\Sets}{\mathbf{Set}}
\newcommand{\blacklater}{\blacktriangleright}
\newcommand{\tot}{\mathcal{S}}
\newcommand{\PSh}{\ensuremath{\textbf{PSh}(\omega)}}

@@

@@

\subsection{Course-of-Value Recursion}
First we have to assume a distributive law $\lambda : DG \to GD$.
For an algebra $a : D(D_{\infty} A) \to A$,
\begin{align*}
  & g : \mu F \to A\\
  & g = a \comp D(G_{\infty}(g) \comp \anacofree{\cata{G^{\lambda}(\InIso)}}) \comp \InOp
\end{align*}
where $G^{\lambda} (alg : D\mu F \to \mu F) = DGX \xrightarrow{\lambda} GDX \xrightarrow{G(alg)} GX  $ is a $D$-algebra on $GX$. The catamorphism of this algebra has the following type
\[
  \cata{G^{\lambda}(\InIso)} : \mu D \to G\mu D
\]
which is the type of a $G$-algebra used to initialise the memory table.

@@

>>>>

\hypersetup{
  linktocpage,
  colorlinks,
  citecolor=BlueViolet,
  filecolor=red,
  linkcolor=Blue,
  urlcolor=BrickRed
}
\newcommand{\mpav}[1]{\textcolor{red}{\textsc{Marco}: #1}}
\newcommand{\dcas}[1]{\textcolor{red}{\textsc{David}: #1}}
\newcommand{\proofcomment}[1]{\text{\{ #1 \}}}
\newenvironment{proofof}[1] {\begin{proof}[Proof of {#1}]}{\end{proof}}
\newcommand{\eqdef}{\stackrel{\mathrm{\Delta}}{=}}
\newcommand{\bnfeq}{\mathrel{::=}}
\newcommand{\defeq}{\triangleq}
\newcommand{\rul}[3]{\frac{#2}{#3}\;  {\textrulelabel{#1}}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\jud}[3]{#1 \vdash #2 : #3}
\newcommand{\bden}[1]{\llparenthesis#1 \rrparenthesis}
\newcommand{\bigslant}[2]{{\raisebox{.2em}{$#1$}\left/\raisebox{-.2em}{$#2$}\right.}}
\newcommand{\quotient}[2]{\bigslant{#1}{#2}}
\newcommand{\curry}{\Lambda}
\newcommand{\uncurry}{\begin{sideways}\begin{sideways}$\Lambda$\end{sideways}\end{sideways}}
\newcommand{\Bool}{\mathbb{B}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Nat}{\N}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Sets}{\mathbf{Set}}
\newcommand{\blacklater}{\blacktriangleright}
\newcommand{\tot}{\mathcal{S}}
\newcommand{\PSh}{\ensuremath{\textbf{PSh}(\omega)}}

@@

\paragraph{Coq Implementation}
Given the two algebras  $a_{1} : D(A_{1} \times A_{2}) \to A_{1}$ and
$a_{2} : D(A_{1} \times A_{2}) \to A_{2}$.
We define the two recursive functions  $f_{1}$ and $f_{2}$ as

\[
  f_{1} = \pi_{1} \comp \hylo{\InIso}{\pair{a_{1}}{a_{2}}} \qquad f_{2} = \pi_{2} \comp \hylo{\InIso}{\pair{a_{1}}{a_{2}}}
\]

@@

\subsection{Course-of-Value Recursion}
First we have to assume a distributive law $\lambda : DG \to GD$.
For an algebra $a : D(D_{\infty} A) \to A$,
\begin{align*}
  & g : \mu D \to A\\
  & g = a \comp D(G_{\infty}(g) \comp \anacofree{\cata{G^{\lambda}(\InIso)}}) \comp \InOp
\end{align*}
where $G^{\lambda} (alg : D\mu F \to \mu F) = DGX \xrightarrow{\lambda} GDX \xrightarrow{G(alg)} GX  $ is a $D$-algebra on $GX$. The catamorphism of this algebra has the following type
\[
  \cata{G^{\lambda}(\InIso)} : \mu D \to G\mu D
\]
which is the type of a $G$-algebra used to initialise the memory table.

@@

\subsubsection{Histomorphism}
The knapsack algorithm
\begin{align*}
  & \operator{knapsack} : [(\Nat, \R)] \to \Nat \to \R\\
  & \operator{knapsack}\; c = \operator{maximum}_{0} [v + \operator{knapsack}\;wvs\; (c - w) \mid (w,v) \leftarrow wvs, w \le c]
\end{align*}

Optimisation of the knapsack with Dynamic Programming.
\begin{align*}
  & \operator{knapsack} : [(\Nat, \R)] \to \Nat \to \R\\
  & \operator{knapsack}\;wvs\;c= table !! c \\
  & \qquad table  = [ks\;i \mid i \leftarrow [0..c]]\\
  & \qquad ks\; i = maximum_{0} [v + table!! (i-w) \mid (w,v) \leftarrow wvs, w \le i]
\end{align*}

\paragraph{Histomorphism}

\begin{align*}
  & \operator{dyna} : [(\Nat, \R)] \to \Nat \to \R\\
  & \operator{knapsack}\;wvs\;c= table !! c\\
  & \qquad table  = [ks\;i \mid i \leftarrow [0..c]]\\
  & \qquad ks\; i = maximum_{0} [v + table!! (i-w) \mid (w,v) \leftarrow wvs, w \le i]
\end{align*}
